---
title: "Archvale Data Analysis"
author: "Nianyu Li"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

------------------------------------------------------------------------

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false

library(tidyverse)
library(ggplot2)
library(skimr)
library(here)
library(lubridate)
library(janitor)
```

# Loading the data

```{r}
#| echo: false
#| message: false
#| warning: false

patient <- read_csv(here::here("data", "patientRecord.csv")) %>% 
  clean_names()
glimpse(patient)

appointment <- read_csv(here::here("data", "appointmentRecord.csv")) %>% 
  clean_names()
glimpse(appointment)

chronicDisease <- read_csv(here::here("data", "chronicDiseaseRecord.csv")) %>% 
  clean_names()
glimpse(chronicDisease)
```

# Data Preparation

Sum the number of consultations of each patients and join the table with `patient` table.

```{r}
#Count visit times of each patients
visit_n <- 
  appointment %>% 
  group_by(patient_id) %>%
  summarise(visit_n = n())

#Join the table with patients record
patients_n_visits <- 
  left_join(x = patient,
            y = visit_n,
            by = "patient_id") %>% 
  
  #change NA to 0 for the patients who have never visited GP
  mutate(visit_n = replace_na(visit_n,0)) %>% 
  
  arrange(desc(visit_n)) %>% 
  
  print()
```

Calculate the patient's GP frequencies to identify the frequent patients.

```{r}
patients_frequency <- patients_n_visits %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()
```

From the cumulative percentages, we can tell that patients who visit the GP more than 7 times accounts for only about 10% of total patients, but more than 45% of the consultations. Therefore, we name this group as `Frequent` and the other as `Non-frequent`.

```{r}
patients_frequency %>%

  #summarize percentage of visits of frequent and non-frequent patients

  mutate(patient_type = case_when(visits_n >= 6 ~ "Frequent(>=6)",
                                  visits_n < 6 ~ "Non-frequent(<6)")) %>%

  group_by(patient_type) %>%

  summarise(pct_visit_total = sum(pct_visit),
            pct_patient_total = sum(pct_patient)) %>%

  ggplot()+

  aes(x = "",
      y = pct_visit_total,
      fill = patient_type) +

  geom_col() +

  scale_fill_brewer()+

  geom_text(aes(label = paste0(pct_visit_total,"%")),
            position = position_stack(vjust = 0.5)) +

  coord_polar(theta = "y")+

  theme_void()+

  theme(axis.ticks = element_blank(),
        axis.text = element_blank())+

  labs(title = "% of Consultation by Patient Type",
       x = NULL, y = NULL,
       fill = "Patient Type")

```

