---
title: "Archvale Data Analysis"
author: "Nianyu Li"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

------------------------------------------------------------------------

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false

library(tidyverse)
library(ggplot2)
library(skimr)
library(here)
library(lubridate)
library(janitor)
library(openxlsx)
library(waterfalls)
```

# Loading the data

```{r}
#| echo: false
#| message: false
#| warning: false

patient <- read_csv(here::here("data", "patientRecord.csv")) %>% 
  clean_names()
glimpse(patient)

appointment <- read_csv(here::here("data", "appointmentRecord.csv")) %>% 
  clean_names()
glimpse(appointment)

chronicDisease <- read_csv(here::here("data", "chronicDiseaseRecord.csv")) %>% 
  clean_names()
glimpse(chronicDisease)
```

# Data Preparation

## Frequencies

Sum the number of consultations of each patients and join the table with `patient` table.

```{r}
#Count visit times of each patients
visit_n <- 
  appointment %>% 
  group_by(patient_id) %>%
  summarise(visit_n = n())

#Join the table with patients record
patients_n_visits <- 
  left_join(x = patient,
            y = visit_n,
            by = "patient_id") %>% 
  
  #change NA to 0 for the patients who have never visited GP
  mutate(visit_n = replace_na(visit_n,0)) %>% 
  
  arrange(desc(visit_n)) %>% 
  
  print()
```

## Age
The age of the patient might affect their behaviour. As the time frame of the appoint record is not that long (only includes record of 2022-2023), we could assume the patient's age is this year minus the year of their birthday, though not 100% accurate.

```{r}
patients_n_visits_1 <-  patients_n_visits %>% 
  
  #convert character to date
  mutate(date_of_birth = dmy(date_of_birth)) %>% 
  
  #calculate age
  mutate(age = year(today()) - year(date_of_birth)) %>% 
  
  print()
```

## Registration Cohort
Patients who registered after March 2023 might have a different behaviour due to the implementation of a marketing campign. Create a new variable to show the registration cohort of the patients: people who registered before March 2023 are classified as in the `Old` cohort, whereas after March 2023 belongs to `New` cohort.

```{r}
patients_n_visits_2 <-  patients_n_visits_1 %>% 
  
  #convert character to date
  mutate(registration_date = dmy(registration_date)) %>%
  
  #create a new variable to show their registration cohort
  mutate(cohort = case_when(registration_date > "2023-03-01" ~ "New",
                            registration_date < "2023-03-01" ~ "Old")) %>% 
  
  print()
```

## Chronic Disease
Patients with chronic diseases visits GP regularly. Create a new variable called chronic_disease to show the number of disease the patient have.

```{r}
#convert the data to long format
chronicDisease_1 <- chronicDisease %>% 
  pivot_longer(cols = -"patient_id",
               names_to = "disease",
               values_to = "yn") %>%
  
  group_by(patient_id) %>% 
  summarise(n_disease = sum(yn))

patients_n_visits_3 <-
  
  #Join the chronic disease table with patients table
  left_join(x = patients_n_visits_2,
            y = chronicDisease_1,
            by = "patient_id") %>% 
  
  #convert NA to 0
  mutate(n_disease = replace_na(n_disease,0)) %>% 
  
  print()

```


# Data Analysis

## Identify Frequent Patients

Calculate the patient's GP frequencies to identify the frequent patients.

```{r}
patients_frequency <- patients_n_visits_3 %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()

#save the tibble as an excel
#write.xlsx(patients_frequency, "patients_frequency.xlsx")
```

From the cumulative percentages, we can tell that patients who visit the GP more than 7 times account for only about 10% of total patients, but more than 45% of the consultations. Therefore, we name this group as `Frequent` and the other as `Non-frequent`.

```{r}
patients_frequency %>%

  #summarize percentage of visits of frequent and non-frequent patients

  mutate(patient_type = case_when(visit_n >= 7 ~ "Frequent(>=7)",
                                  visit_n < 7 ~ "Non-frequent(<7)")) %>%
  
  group_by(patient_type) %>%

  summarise(pct_visit_total = sum(visit_pct),
            pct_patient_total = sum(patient_pct)) %>%
  
  ggplot()+

  aes(x = '',
      y = pct_visit_total,
      fill = patient_type) +

  geom_col(position = "stack") +

  scale_fill_brewer()+

  geom_text(aes(label = paste0(pct_visit_total,"%")),
            position = position_stack(vjust = 0.5)) +

  coord_polar("y")+

  theme_void()+

  theme(axis.ticks = element_blank(),
        axis.text = element_blank())+

  labs(title = "% of Consultation by Patient Type",
       x = NULL, y = NULL,
       fill = "Patient Type")

```



```{r}

```

