---
title: "Archvale Data Analysis"
author: "Nianyu Li"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

------------------------------------------------------------------------

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false

library(tidyverse)
library(ggplot2)
library(skimr)
library(here)
library(lubridate)
library(janitor)
library(openxlsx)
library(waterfalls)
library(GGally)
```

# Loading the data

```{r}
#| echo: false
#| message: false
#| warning: false

patient <- read_csv(here::here("data", "patientRecord.csv")) %>% 
  clean_names()
glimpse(patient)

appointment <- read_csv(here::here("data", "appointmentRecord.csv")) %>% 
  clean_names()
glimpse(appointment)

chronicDisease <- read_csv(here::here("data", "chronicDiseaseRecord.csv")) %>% 
  clean_names()
glimpse(chronicDisease)
```

# Data Preparation

## Frequencies

Sum the number of consultations of each patients and join the table with `patient` table.

```{r}
#Count visit times of each patients
visit_n <- 
  appointment %>% 
  group_by(patient_id) %>%
  summarise(visit_n = n())

#Join the table with patients record
patients_n_visits <- 
  left_join(x = patient,
            y = visit_n,
            by = "patient_id") %>% 
  
  #change NA to 0 for the patients who have never visited GP
  mutate(visit_n = replace_na(visit_n,0)) %>% 
  
  arrange(desc(visit_n)) %>% 
  
  print()
```

## Age

The age of the patient might affect their behaviour. As the time frame of the appoint record is not that long (only includes record of 2022-2023), we could assume the patient's age is this year minus the year of their birthday, though not 100% accurate.

```{r}
patients_n_visits_1 <-  patients_n_visits %>% 
  
  #convert character to date
  mutate(date_of_birth = dmy(date_of_birth)) %>% 
  
  #calculate age
  mutate(age = year(today()) - year(date_of_birth)) %>% 
  
  print()
```

## Registration Cohort

Patients who registered after March 2023 might have a different behaviour due to the implementation of a marketing campign. Create a new variable to show the registration cohort of the patients: people who registered before March 2023 are classified as in the `Old` cohort, whereas after March 2023 belongs to `New` cohort.

```{r}
patients_n_visits_2 <-  patients_n_visits_1 %>% 
  
  #convert character to date
  mutate(registration_date = dmy(registration_date)) %>%
  
  #create a new variable to show their registration cohort
  mutate(cohort = case_when(registration_date > "2023-03-01" ~ "New",
                            registration_date < "2023-03-01" ~ "Old")) %>% 
  
  print()
```

## Chronic Disease

Patients with chronic diseases visits GP regularly. Create a new variable called chronic_disease to show the number of disease the patient have.

```{r}
#convert the data to long format
chronicDisease_1 <- chronicDisease %>% 
  pivot_longer(cols = -"patient_id",
               names_to = "disease",
               values_to = "yn") %>%
  
  group_by(patient_id) %>% 
  summarise(n_disease = sum(yn))

patients_n_visits_3 <-
  
  #Join the chronic disease table with patients table
  left_join(x = patients_n_visits_2,
            y = chronicDisease_1,
            by = "patient_id") %>% 
  
  #convert NA to 0
  mutate(n_disease = replace_na(n_disease,0)) %>% 
  
  print()

```

## Average appointment time

```{r}
#| warning: false

#Calculate the average appointment time for patients
app_time <- appointment %>% 
  group_by(patient_id) %>% 
  summarise(avg_time = mean(as.integer(planned_appointment_duration)))

#join it with the patients table
patients_n_visits_4 <- left_join(x = patients_n_visits_3,
                                 y = app_time,
                                 by = "patient_id") %>% 
  
  #convert NA to 0
  mutate(avg_time = replace_na(avg_time,0)) %>% 
  
  print()
```


## Average appointment intervals

```{r}
app_interval <- appointment %>% 
  
  #convert character to date
  mutate(appointment_date = dmy(appointment_date)) %>%
  
  #group by patients
  group_by(patient_id) %>% 
  
  #calculate average interval
  summarise(app_interval_sum = max(appointment_date) - min(appointment_date),
            visit_n = n()) %>% 
  
  mutate(app_interval = app_interval_sum/visit_n,
         app_interval_int = as.numeric(app_interval)) %>% 
  
  #arrange the data
  arrange(desc(app_interval)) %>%
  
  select(patient_id, app_interval_int)
  
  #Join it with patients table and omit NAs
patients_n_visits_5 <- left_join(x = patients_n_visits_4,
                                 y = app_interval,
                                 by = "patient_id") %>% 
  
  #convert NA to 0
  mutate(app_interval_int = replace_na(app_interval_int,0)) %>% 
  
  print()
```


# Data Analysis

We make the following assumptions to identify the inappropriate users of GP services:

1)  They are frequent users of GP services.

2)  They are associate with potential inappropriate usage.

## Identify Frequent Patients

### General Approach

Calculate the patient's GP frequencies to identify the frequent patients.

```{r}
patients_frequency <- patients_n_visits_5 %>%
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()

#save the tibble as an excel
#write.xlsx(patients_frequency, "patients_frequency.xlsx")
```

From the cumulative percentages, we can tell that patients who visit the GP more than 6 times account for only about 13% of total patients, but more than 50% of the consultations. Therefore, we name this group as `Frequent` and the other as `Non-frequent`.

```{r}
patients_frequency %>%

  #summarize percentage of visits of frequent and non-frequent patients

  mutate(patient_type = case_when(visit_n >= 6 ~ "Frequent(>=6)",
                                  visit_n < 6 ~ "Non-frequent(<6)")) %>%
  
  group_by(patient_type) %>%

  summarise(pct_visit_total = sum(visit_pct),
            pct_patient_total = sum(patient_pct)) %>%
  
  ggplot()+

  aes(x = '',
      y = pct_visit_total,
      fill = patient_type) +

  geom_col(position = "stack") +

  scale_fill_brewer()+

  geom_text(aes(label = paste0(pct_visit_total,"%")),
            position = position_stack(vjust = 0.5)) +

  coord_polar("y")+

  theme_void()+

  theme(axis.ticks = element_blank(),
        axis.text = element_blank())+

  labs(title = "% of Consultation by Patient Type",
       x = NULL, y = NULL,
       fill = "Patient Type")

```

### Testing the hypothesis

Test the hypothesis: does GP frequencies have to do with Age/Gender/Chronic Disease/Registration Cohort. If there is a relationship, we might take that into consideration when we decide whether the patient has an appropriate clinical need.

```{r}
patients_n_visits_5 %>% 
  select(age, n_disease, visit_n) %>% 
  ggpairs()
```

```{r}
patients_n_visits_5 %>% 
  select(gender, cohort, visit_n) %>% 
  ggpairs(aes(fill = gender))+
  scale_fill_manual(values = c("#001e62","#c6cddc"))
```

We can tell that Age and Chronic Disease have significant positive correlations with frequency of consultations. It also varies in Gender and Registration Cohort. Female tends to have higher number of consultations. New cohort tends to have lower number of consultations.

### Refined Approach: Frequent patients by segments

People in different age/gender/registration cohort behave differently. Thus, different criteria is used to identify frequent patients.

#### Gender

```{r}
#Filter data for just female
patients_frequency_f <- patients_n_visits_5 %>%
  
  filter(gender == "Female") %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()
```

```{r}
patients_frequency_f %>%

  #summarize percentage of visits of frequent and non-frequent patients

  mutate(patient_type = case_when(visit_n >= 7 ~ "Frequent(>=7)",
                                  visit_n < 7 ~ "Non-frequent(<7)")) %>%
  
  group_by(patient_type) %>%

  summarise(pct_visit_total = sum(visit_pct),
            pct_patient_total = sum(patient_pct)) %>%
  
  ggplot()+

  aes(x = '',
      y = pct_visit_total,
      fill = patient_type) +

  geom_col(position = "stack") +

  scale_fill_brewer()+

  geom_text(aes(label = paste0(pct_visit_total,"%")),
            position = position_stack(vjust = 0.5)) +

  coord_polar("y")+

  theme_void()+

  theme(axis.ticks = element_blank(),
        axis.text = element_blank())+

  labs(title = "% of Consultation by Patient Type - Female",
       x = NULL, y = NULL,
       fill = "Patient Type")

```


```{r}
#Filter data for just male
patients_frequency_m <- patients_n_visits_5 %>%
  
  filter(gender == "Male") %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()
```

```{r}
patients_frequency_m %>%

  #summarize percentage of visits of frequent and non-frequent patients

  mutate(patient_type = case_when(visit_n >= 6 ~ "Frequent(>=6)",
                                  visit_n < 6 ~ "Non-frequent(<6)")) %>%
  
  group_by(patient_type) %>%

  summarise(pct_visit_total = sum(visit_pct),
            pct_patient_total = sum(patient_pct)) %>%
  
  ggplot()+

  aes(x = '',
      y = pct_visit_total,
      fill = patient_type) +

  geom_col(position = "stack") +

  scale_fill_brewer()+

  geom_text(aes(label = paste0(pct_visit_total,"%")),
            position = position_stack(vjust = 0.5)) +

  coord_polar("y")+

  theme_void()+

  theme(axis.ticks = element_blank(),
        axis.text = element_blank())+

  labs(title = "% of Consultation by Patient Type - Male",
       x = NULL, y = NULL,
       fill = "Patient Type")

```

#### Age group

```{r}
#Assign patients to "Early_childhood(<8) and elderly(>65)" and "Others"

#Create a new variable to label patients
patients_n_visits_5 <-patients_n_visits_5 %>% 
  mutate(age_group = case_when(age %in% (8:65) ~ "Others",
                               !age %in% (8:65) ~ "Under 8 or above 65")) %>% 
  print()
```


```{r}
#Filter data for just "Early_childhood(<8) and elderly(>65)"
patients_frequency_age <- patients_n_visits_5 %>%
  
  filter(age_group == "Under 8 or above 65") %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()
```

```{r}
patients_frequency_age %>%

  #summarize percentage of visits of frequent and non-frequent patients

  mutate(patient_type = case_when(visit_n >= 7 ~ "Frequent(>=7)",
                                  visit_n < 7 ~ "Non-frequent(<7)")) %>%
  
  group_by(patient_type) %>%

  summarise(pct_visit_total = sum(visit_pct),
            pct_patient_total = sum(patient_pct)) %>%
  
  ggplot()+

  aes(x = '',
      y = pct_visit_total,
      fill = patient_type) +

  geom_col(position = "stack") +

  scale_fill_brewer()+

  geom_text(aes(label = paste0(pct_visit_total,"%")),
            position = position_stack(vjust = 0.5)) +

  coord_polar("y")+

  theme_void()+

  theme(axis.ticks = element_blank(),
        axis.text = element_blank())+

  labs(title = "% of Consultation by Patient Type - Under 8 or above 65",
       x = NULL, y = NULL,
       fill = "Patient Type")

```

```{r}
#Filter data for just "others"
patients_frequency_age_o <- patients_n_visits_5 %>%
  
  filter(age_group == "Others") %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()
```

```{r}
patients_frequency_age_o %>%

  #summarize percentage of visits of frequent and non-frequent patients

  mutate(patient_type = case_when(visit_n >= 6 ~ "Frequent(>=6)",
                                  visit_n < 6 ~ "Non-frequent(<6)")) %>%
  
  group_by(patient_type) %>%

  summarise(pct_visit_total = sum(visit_pct),
            pct_patient_total = sum(patient_pct)) %>%
  
  ggplot()+

  aes(x = '',
      y = pct_visit_total,
      fill = patient_type) +

  geom_col(position = "stack") +

  scale_fill_brewer()+

  geom_text(aes(label = paste0(pct_visit_total,"%")),
            position = position_stack(vjust = 0.5)) +

  coord_polar("y")+

  theme_void()+

  theme(axis.ticks = element_blank(),
        axis.text = element_blank())+

  labs(title = "% of Consultation by Patient Type - Others",
       x = NULL, y = NULL,
       fill = "Patient Type")

```

#### Chronic Disease

```{r}
#Filter data for just people with chronic disease
patients_frequency_disease <- patients_n_visits_5 %>%
  
  filter(n_disease > 0) %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()
```

```{r}
patients_frequency_disease %>%

  #summarize percentage of visits of frequent and non-frequent patients

  mutate(patient_type = case_when(visit_n >= 7 ~ "Frequent(>=7)",
                                  visit_n < 7 ~ "Non-frequent(<7)")) %>%
  
  group_by(patient_type) %>%

  summarise(pct_visit_total = sum(visit_pct),
            pct_patient_total = sum(patient_pct)) %>%
  
  ggplot()+

  aes(x = '',
      y = pct_visit_total,
      fill = patient_type) +

  geom_col(position = "stack") +

  scale_fill_brewer()+

  geom_text(aes(label = paste0(pct_visit_total,"%")),
            position = position_stack(vjust = 0.5)) +

  coord_polar("y")+

  theme_void()+

  theme(axis.ticks = element_blank(),
        axis.text = element_blank())+

  labs(title = "% of Consultation by Patient Type - With Chronic Disease",
       x = NULL, y = NULL,
       fill = "Patient Type")

```

```{r}
#Filter data for just people without chronic disease
patients_frequency_disease <- patients_n_visits_5 %>%
  
  filter(n_disease == 0) %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()
```
```{r}
patients_frequency_disease %>%

  #summarize percentage of visits of frequent and non-frequent patients

  mutate(patient_type = case_when(visit_n >= 6 ~ "Frequent(>=6)",
                                  visit_n < 6 ~ "Non-frequent(<6)")) %>%
  
  group_by(patient_type) %>%

  summarise(pct_visit_total = sum(visit_pct),
            pct_patient_total = sum(patient_pct)) %>%
  
  ggplot()+

  aes(x = '',
      y = pct_visit_total,
      fill = patient_type) +

  geom_col(position = "stack") +

  scale_fill_brewer()+

  geom_text(aes(label = paste0(pct_visit_total,"%")),
            position = position_stack(vjust = 0.5)) +

  coord_polar("y")+

  theme_void()+

  theme(axis.ticks = element_blank(),
        axis.text = element_blank())+

  labs(title = "% of Consultation by Patient Type - Without Chronic Disease",
       x = NULL, y = NULL,
       fill = "Patient Type")

```

#### Registration Cohort

```{r}
#Filter data for just people in the "New" cohort
patients_frequency_cohort <- patients_n_visits_5 %>%
  
  filter(cohort == "New") %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()
```

```{r}
#Filter data for just people in the "New" cohort
patients_frequency_cohort_old <- patients_n_visits_5 %>%
  
  filter(cohort == "Old") %>% 
  
  #group the patients by their frequencies
  group_by(visit_n) %>% 
  
  #count the total number of patients and visits in each group
  summarise(patient_n = n(),
            visit_total_n = sum(visit_n)) %>% 
  
  #calculate the percentages
  mutate(patient_pct = round(patient_n/sum(patient_n)*100,2),
         visit_pct = round(visit_total_n/sum(visit_total_n)*100,2)) %>% 
  
  #calculate the cumulative percentages
  arrange(desc(visit_n)) %>% 
  mutate(patient_cumpct = cumsum(patient_pct),
         visit_cumpct = cumsum(visit_pct)) %>% 
  
  print()
```

As we can see, new cohort does not appears to behave as the 10%-50% pattern we discussed before, and 93% of the patients haven't visited GP, therefore, we only filter for old cohort to find inappropriate usage.

## Filter for frequent patient

```{r}
patients_oldco <- patients_n_visits_5 %>% 
  
  #Filter only for old cohort
  filter(cohort == "Old") %>% 
  
  print()
```
Now we filter for frequent patients in vulnerable groups, using the 7-times-visit criteria.
```{r}
patients_oldco_vul <- patients_oldco %>% 
  
  #filter vulnerable groups 
  filter(gender == "Female" 
  | age_group == "Under 8 or above 65"
  | n_disease > 0) %>% 
  
  #filter for frequent patient
  filter(visit_n >= 7) %>% 
  
  print()
```
Now we filter for frequent patients in none-vulnerable groups, using the 6-times-visit criteria.

```{r}
patients_oldco_nonevul <- patients_oldco %>% 
  
  filter(gender == "Male" 
  & age_group == "Others"
  & n_disease == 0) %>% 
  
  #filter for frequent patient
  filter(visit_n >= 6) %>% 
  
  print()
```
We have a total of 142+22=164 patients. We now look at their appointment patterns to find the top 50 that are most likely to use GP inappropriately.

## Identify inappropriate usage

Filter out the patients using the following features.

    Features include:

-   Shorter average planned time;
-   Shorter Intervals between appointments;

```{r}
# first for vulnerable patients
patients_oldco_vul_sus <- patients_oldco_vul %>%
  
  # Shorter average planned time
  filter(avg_time<quantile(avg_time, 0.5) | 
         # Shorter Intervals between appointments
         app_interval_int<quantile(app_interval_int, 0.5)) %>% 
  
  print()
```
Usually, the most vulnerable groups (people in under 8 and above 65 age groups and with chronic diseases) in are less likely to use GP inappropriately. So we try to filter them out.

```{r}
patients_oldco_vul_sus_1 <- patients_oldco_vul_sus %>% 
  filter(age_group != "Under 8 or above 65" & n_disease == 0) %>% 
  print()
```

```{r}
# then for none-vulnerable patients
patients_oldco_nonevul_sus <- patients_oldco_nonevul %>%
  
  # Shorter average planned time
  filter(avg_time<quantile(avg_time, 0.5) | 
         # Shorter Intervals between appointments
         app_interval_int<quantile(app_interval_int, 0.5)) %>% 
  
  print()
```

Now we have a total of 29+17=46 patients. We merge the two tables to get the top patients that are likely to use GP appropriately.

```{r}
final_patient_list <- rbind(patients_oldco_nonevul_sus,
                            patients_oldco_vul_sus_1) %>% 
  print()

#save the tibble as an excel
#write.xlsx(final_patient_list, "final_patient_list.xlsx")

```
Filter out the patients who are:
 1) under 8 and above 65
 2) with chronic disease
 3) in the new cohort
```{r}
patients_test <- patients_n_visits_5 %>% 
  filter(age_group == "Others"
         & n_disease == 0
         & cohort == "Old") %>% 
  print()
```



